---
title: "cwd.datacleaning.updated.Rmd"
author: "Kaija Gahm"
date: "10/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load functions and libraries
```{r results = "hide"}
source("ownfunctions.R")
source("libraries.R")
```

# _______________________
# Upper River (4, 8, 13)
## Load data
```{r}
load(here("Analysis", "data", "outputs", "fish_data_EF.Rda"))
head(fish_data_EF)
  # Fish sampling data
fish_data_EF <- fish_data_EF %>% mutate(fishdata_rownum = 1:nrow(fish_data_EF)) #number the rows
sites_p4p8p13 <- read.csv(here("Analysis", "data", "inputs", "DataSets_7_7_upper", "AttributeTables", "sites_p4p8p13.txt")) 
  # reprojected data
rivmi <- read.table(here("Analysis", "data", "inputs", "p4p8p13_rivmile.txt"), sep = ",", header = T) 
  # river mile data
sites_aa_5m <- read.csv(here("Analysis", "data", "inputs", "DataSets_7_7_upper", "AttributeTables", "sites_aquaareas5m.txt"))# merged aquatic areas data for points buffered by 5 meters.

## Environmental data
sites_terrestrial <- read.table(here("Analysis", "data", "inputs", "DataSets_7_7_upper", "AttributeTables", "sites_terrestrial.txt"), header = T, sep = ",") # nearest terrestrial area info for each point
terrestrial_forests <- read.table(here("Analysis", "data", "inputs", "DataSets_7_7_upper", "AttributeTables", "Terrestrial_Forests.txt"), sep = ",", header = T)# info about terrestrial areas/forests
sites_forest <- read.table(here("Analysis", "data", "inputs", "DataSets_7_7_upper", "AttributeTables", "sites_forest.txt"), sep = ",", header = T) # nearest forest info for each point
lc_2010 <- read.table(here("Analysis", "data", "inputs", "DataSets_7_7_upper", "AttributeTables", "lc_2010.txt"), sep = ",", header = T)# 2010 landcover info
```

## Clean fish data
```{r}
fish_data_EF <- fish_data_EF %>%
  filter(sitetype %in% c(0, 1), # remove sitetype 2
         gear == "D") # remove fyke net (we want only D = daytime electrofishing)
```

## Initialize the data frame
```{r}
sites_ur <- rivmi %>% dplyr::select(TARGET_FID, Field1, stratum, lcode, sdate, utm_e, utm_n, snag, RIVER_MILE)
```

## Join fish data
```{r}
sites_ur <- sites_ur %>% rename(fishdata_rownum = Field1) # The Field1 column in this dataset corresponds to the row number in fish_data_EF.
fish_data_EF_tojoin <- fish_data_EF %>% dplyr::select(sitetype, gear, wingdyke, riprap, trib, barcode, summary, fishdata_rownum, zone15e, zone15n, utmzone)
sites_ur <- left_join(sites_ur, fish_data_EF_tojoin, by = "fishdata_rownum")
# Note that after this join, we now have the "barcode" column in our data. This will allow us to join other data by barcode.
sites_ur %$% table(utmzone)
```

## Join aquatic areas data
```{r}
# Make unique column for joining in each
sites_aa_5m <- sites_aa_5m %>% mutate(joiner = paste(stratum, lcode, sdate, utm_e, utm_n, snag))
sites_ur <- sites_ur %>% mutate(joiner = paste(stratum, lcode, sdate, utm_e, utm_n, snag))
# These appear to be exactly the same. Check:
sum(sites_aa_5m$joiner != sites_ur$joiner) #yep, they are the same! That makes it easy:
sites_ur <- sites_ur %>% dplyr::select(-c(joiner, TARGET_FID)) # remove the joiner column and also TARGET_FID; we don't need that anymore.
sites_aa_5m <- sites_aa_5m %>% dplyr::select(-c(joiner, TARGET_FID)) 
# Join the aa information to sites_ur.
sites_ur <- left_join(sites_ur, sites_aa_5m, by = c("stratum", "lcode", "sdate", "utm_e", "utm_n", "snag"))
```

## Clean aquatic areas
```{r}
# "Observations with value of 0 in all the columns from aqa_2010_lvl3_011918.shp do not intersect with the aquatic areas layer". 
# remove sites that are NA for Perimeter, Area, and avg_fetch (i.e. don't intersect the aquatic areas layer)
sites_ur <- sites_ur %>%
  filter(!(Perimeter == 0 & Area == 0 & avg_fetch == 0))
```

```{r}
# Fix missing pools
# there are a concerning number of NA's in the `pool` column that shouldn't be there. Luckily, the `uniq_id` column tells us which pool these are from. 
addpools <- function(df){
  pools <- as.numeric(substr(x = as.character(df$uniq_id), 
                             start = 2, 
                             stop = 3))
  df$pool[is.na(df$pool)] <- pools[is.na(df$pool)]
  return(df)
}
sites_ur <- addpools(sites_ur)
locate.nas(sites_ur) # we still have some NA's for "snag", but that is to be expected. There are also some NA's for `pool`, but those have NA for snag and have some sketchy-looking barcodes.
# More NA pools fixing
sum(is.na(sites_ur$pool)) # we still have a bunch of NA's for pool
# Make a plot to see what's going on
sites_ur %>% ggplot(aes(x = utm_e, y = utm_n))+
  geom_point(aes(col = pool)) # This shows some pretty clear divisions between pools along the river. The NA's are evenly mixed in with the other points.
# Assign the NA pools based on lat/lon
napools <- sites_ur %>% filter(is.na(pool))
napools <- napools %>% mutate(pool = case_when(utm_n > 4900000 ~ "P04",
                                               utm_n > 4800000 & utm_n < 4900000 ~ "P08",
                                               utm_n < 4700000 ~ "P13"))
sites_ur$pool[is.na(sites_ur$pool)] <- napools$pool
# Double check with a plot
sites_ur %>% ggplot(aes(x = utm_e, y = utm_n))+
  geom_point(aes(col = pool)) # Looks good!
```


<!-- ## Join environmental info -->
<!-- ### Distance to terrestrial areas -->
<!-- ```{r} -->
<!-- # terrestrial data to join -->
<!-- terr_tojoin <- sites_terrestrial %>% dplyr::select(NEAR_FID, NEAR_DIST, Field1) -->
<!-- sites_ur <- left_join(sites_ur, terr_tojoin, by = "Field1") %>%  -->
<!--   rename(NEAR_TERR_FID = NEAR_FID, # ID of nearest terrestrial region -->
<!--          NEAR_TERR_DIST = NEAR_DIST) # distance to nearest terrestrial region -->

<!-- # The is the join that I think I was trying to do originally. It assumes that Field1 in sites_terrestrial is equivalent to Field1 in sites_aa_5m (and therefore sites_ur). Need to check that this is right. -->

<!-- # ID of nearest terrestrial region -->
<!-- ``` -->

<!-- ### Info about nearest terrestrial region -->
<!-- ```{r} -->
<!-- # We're going to pull terrestrial areas information columns from `lc_2010`, not from `terrestrial`, because the FID's don't match up in `terrestrial`. -->
<!-- terrinfo_tojoin <- lc_2010 %>% dplyr::select(FID, CLASS_7_C, CLASS_7_N) -->
<!-- sites_ur <- left_join(sites_ur, terrinfo_tojoin, by = c("NEAR_TERR_FID" = "FID")) %>% -->
<!--   rename(NEAR_TERR_CLASS_7 = CLASS_7_C, -->
<!--          NEAR_TERR_CLASS_7_N = CLASS_7_N) # join the columns and rename appropriately -->
<!-- ``` -->

<!-- ### Distance to forested areas -->
<!-- ```{r} -->
<!-- forest_tojoin <- sites_forest %>% dplyr::select(Field1, NEAR_FID, NEAR_DIST) -->
<!-- sites_ur <- left_join(sites_ur, forest_tojoin, by = "Field1") %>%  -->
<!--   rename(NEAR_FOREST_FID = NEAR_FID, # which forested area is the closest -->
<!--          NEAR_FOREST_DIST = NEAR_DIST) # distance to nearest forested area -->
<!-- ``` -->

## Cleaning
### Add a year column
```{r}
# Add a `year` column
sites_ur <- sites_ur %>% mutate(year = str_extract(sdate, "[[:digit:]]{4}"))
```

### Select variables
```{r}
toinclude <- c("fishdata_rownum", "stratum", "lcode", "sdate", "utm_e", "utm_n", "snag", "RIVER_MILE", "wingdyke", "riprap", "barcode", "summary", "zone15e", "zone15n", "utmzone", "uniq_id", "pool", "Perimeter", "max_depth", "avg_depth", "tot_vol", "sdi", "pct_terr", "pct1wetf", "year")
sites_ur <- sites_ur %>% dplyr::select(all_of(toinclude)) # choose which columns to include
names(sites_ur)
```

### Change some names
```{r}
# Change some names
sites_ur <- sites_ur %>% rename(perimeter = Perimeter,
                                      shoreline_density_index = sdi,
                                      pct_prm_wetf = pct1wetf,
                                      river_mile = RIVER_MILE)
```

### Replace -999 etc. with NA
```{r}
factors <- sites_ur %>% dplyr::select_if(., is.factor) %>% names()
colstomodify <- sites_ur %>% dplyr::select(-all_of(factors)) %>% dplyr::select(-barcode) %>% names()
               
for(i in colstomodify){
  sites_ur[,i][sites_ur[,i] < -5000] <- NA
}                                   
```

### Investigate NA's
```{r}
locate.nas(sites_ur)
# These seem fine
```

### Remove sites NA for snag
```{r}
sites_ur <- sites_ur %>% filter(!is.na(snag))
```

### Exclude UXO and TWZ
```{r}
sites_ur <- sites_ur %>% filter(!(stratum %in% c("UXO", "TWZ")))
```

### Remove pseudo-shoreline observations
```{r}
ps_ur <- sites_ur %>% filter(summary == "7")
ps_ur %$% table(stratum, pool)
sites_ur <- sites_ur %>% filter(summary != "7") # a code of 7 in the summary column means that the observations are pseudo-shoreline and should be excluded.
```

### Check NA's again
```{r}
locate.nas(sites_ur)
```

## Save data
```{r}
#save(sites_ur, file = "../data/outputs/sites_ur.Rda")
```

# _______________________
# Lower River (26, La Grange, Open River)
## Load data
```{r}
# Load files
# Fish sampling data - made by MVA 10/31/18 by importing csv "ltrm_fish_data.csv" and exporting as Rda. 11/26/19: KJGG could not find this file. Re-imported ltrm_fish_data.csv and saved as .Rda.
# ltrm_fish_data_new <- read.csv(here("Analysis", "data", "inputs", "DataSets_8_17_lower", "ltrm_fish_data.csv"), header = T)
#  save(ltrm_fish_data_new, file=here("Analysis", "data", "inputs", "DataSets_8_17_lower", "ltrm_fish_data_new.Rda"))

load(here("Analysis", "data", "inputs", "DataSets_8_17_lower", "ltrm_fish_data_new.Rda")) # This is the equivalent of fish_data_EF for the upper river.

ltrm_fish_data <- ltrm_fish_data %>% mutate(fishdata_rownum = 1:nrow(ltrm_fish_data)) #number the rows

sites_p26ilror <- read.table(here("Analysis", "data", "inputs", "DataSets_8_17_lower", "AttributeTables", "sites_p26ilror.txt"), header = T, sep = ",") # reprojected data. This is the equivalent of sites_p4p8p13.
# I can't find an equivalent of the river miles file for the lower river. Do we have?

sites_aa_5m <- read.table(here("Analysis", "data", "inputs", "DataSets_8_17_lower", "AttributeTables", "sites_aquaareas5m2.txt"), header = T, sep = ",") # merged aquatic areas data for points buffered by 5 meters.

## Environmental data
sites_terrestrial <- read.table(here("Analysis", "data", "inputs", "DataSets_8_17_lower", "AttributeTables", "sites_terrestrial2.txt"), header = T, sep = ",") # nearest terrestrial area info for each point
terrestrial_forests <- read.table(here("Analysis", "data", "inputs", "DataSets_8_17_lower", "AttributeTables", "Terrestrial_Forests2.txt"), header = T, sep = ",") # info about terrestrial areas/forests
sites_forest <- read.table(here("Analysis", "data", "inputs", "DataSets_8_17_lower", "AttributeTables", "sites_forest2.txt"), header = T, sep = ",") # nearest forest info for each point
lc_2010 <- read.table(here("Analysis", "data", "inputs", "DataSets_8_17_lower", "AttributeTables", "lc_20102.txt"), header = T, sep = ",") # landcover info
```

## Clean fish data
```{r}
ltrm_fish_data <- ltrm_fish_data %>%
  filter(sitetype %in% c(0, 1), # remove sitetype 2
         gear == "D") # remove fyke net (we want only D = daytime electrofishing)
```

## Initialize the data frame
```{r}
sites_lr <- sites_p26ilror
```

## Join fish data
```{r}
sites_lr <- sites_lr %>% rename(fishdata_rownum = RowID_) # The RowID_ column in this dataset corresponds to the row number in ltrm_fish_data, I think. I'm inferring this from parallels to the upper river data.
ltrm_fish_data_tojoin <- ltrm_fish_data %>% dplyr::select(sitetype, gear, wingdyke, riprap, trib, barcode, summary, fishdata_rownum, zone15e, zone15n, utmzone)
sites_lr <- left_join(sites_lr, ltrm_fish_data_tojoin, by = "fishdata_rownum")
# Note that after this join, we now have the "barcode" column in our data. This will allow us to join other data by barcode.
```

## Join aquatic areas data
```{r}
# Make unique column for joining in each
sites_aa_5m <- sites_aa_5m %>% mutate(joiner = paste(stratum, lcode, sdate, utm_e, utm_n, snag))
sites_lr <- sites_lr %>% mutate(joiner = paste(stratum, lcode, sdate, utm_e, utm_n, snag))
# These appear to be exactly the same. Check:
sum(sites_aa_5m$joiner != sites_lr$joiner) #yep, they are the same! That makes it easy:
nrow(sites_aa_5m) == nrow(sites_lr) #TRUE
sites_lr <- sites_lr %>% dplyr::select(-c(joiner, FID)) # remove the joiner column and also FID; we don't need that anymore.
sites_aa_5m <- sites_aa_5m %>% dplyr::select(-c(joiner, FID))
# Join the aa information to sites_lr.
sites_lr <- left_join(sites_lr, sites_aa_5m, by = c("stratum", "lcode", "sdate", "utm_e", "utm_n", "snag"))
```

## Clean aquatic areas
```{r}
# "Observations with value of 0 in all the columns from aqa_2010_lvl3_011918.shp do not intersect with the aquatic areas layer". 
sites_lr <- sites_lr %>% # remove observations that are missing values for several environmental vars
  filter(!(Perimeter == 0 & Area == 0 & avg_fetch == 0))
```

## Fix pools
```{r}
# Fix missing pools using addpools function defined for ur above.
sites_lr <- addpools(sites_lr)
locate.nas(sites_lr) # Fewer NA's for `snag` than for the upper river data set, but there are many NA's for the environmental data. I wonder if these are problematic?
# More NA pools fixing
sum(is.na(sites_lr$pool)) # we still have a bunch of NA's for pool
```

## Investigate problematic utm coords
```{r}
# Make a plot to see what's going on
sites_lr %>% ggplot(aes(x = utm_e, y = utm_n))+
  geom_point(aes(col = pool)) # There are some very weird things going on with the UTM coordinates here.
# Investigate the weird utm coordinates
sites_lr %$% table(utmzone) # indeed, these are listed as being in zone 16. 
# But if we plot the zone 15 coords, it seems to be fine? (except for a few points)
sites_lr %>% ggplot(aes(x = zone15e, y = zone15n))+
  geom_point(aes(col = pool))
```

<!-- ## Join environmental info -->
<!-- ### Distance to terrestrial areas -->
<!-- ```{r} -->
<!-- # terrestrial data to join -->
<!-- terr_tojoin <- sites_terrestrial %>% dplyr::select(NEAR_FID, NEAR_DIST, RowID_) -->
<!-- sites_lr <- left_join(sites_lr, terr_tojoin, by = "RowID_") %>%  -->
<!--   rename(NEAR_TERR_FID = NEAR_FID, # ID of nearest terrestrial region -->
<!--          NEAR_TERR_DIST = NEAR_DIST) # distance to nearest terrestrial region -->

<!-- # The is the join that I think I was trying to do originally. It assumes that RowID_ in sites_terrestrial is equivalent to RowID_ in sites_aa_5m (and therefore sites_lr). Need to check that this is right. -->

<!-- # ID of nearest terrestrial region -->
<!-- ``` -->

<!-- ### Info about nearest terrestrial region -->
<!-- ```{r} -->
<!-- # We're going to pull terrestrial areas information columns from `lc_2010`, not from `terrestrial`, because the FID's don't match up in `terrestrial`. -->
<!-- terrinfo_tojoin <- lc_2010 %>% dplyr::select(FID, CLASS_7_C, CLASS_7_N) -->
<!-- sites_lr <- left_join(sites_lr, terrinfo_tojoin, by = c("NEAR_TERR_FID" = "FID")) %>% -->
<!--   rename(NEAR_TERR_CLASS_7 = CLASS_7_C, -->
<!--          NEAR_TERR_CLASS_7_N = CLASS_7_N) # join the columns and rename appropriately -->
<!-- ``` -->

<!-- ### Distance to forested areas -->
<!-- ```{r} -->
<!-- forest_tojoin <- sites_forest %>% dplyr::select(RowID_, NEAR_FID, NEAR_DIST) -->
<!-- sites_lr <- left_join(sites_lr, forest_tojoin, by = "RowID_") %>%  -->
<!--   rename(NEAR_FOREST_FID = NEAR_FID, # which forested area is the closest -->
<!--          NEAR_FOREST_DIST = NEAR_DIST) # distance to nearest forested area -->
<!-- ``` -->

## Cleaning
### Add a year column
```{r}
# Add a `year` column
sites_lr <- sites_lr %>% mutate(year = str_extract(sdate, "[[:digit:]]{4}"))
```

### Select variables
```{r}
toinclude <- c("fishdata_rownum", "stratum", "lcode", "sdate", "utm_e", "utm_n", "snag", "wingdyke", "riprap", "barcode", "summary", "zone15e", "zone15n", "utmzone", "OBJECTID_1", "RowID_", "uniq_id", "pool", "Perimeter", "max_depth", "avg_depth", "tot_vol", "sdi", "pct_terr", "pct1wetf", "year")
sites_lr <- sites_lr %>% dplyr::select(all_of(toinclude)) # choose which columns to include
names(sites_lr)
```

### Change some names
```{r}
# Change some names
sites_lr <- sites_lr %>% rename(perimeter = Perimeter,
                                      shoreline_density_index = sdi,
                                      pct_prm_wetf = pct1wetf) # would need to rename river mile if it were in here.
```

### Replace -999 etc. with NA
```{r}
factors <- sites_lr %>% dplyr::select_if(., is.factor) %>% names()
colstomodify <- sites_lr %>% dplyr::select(-factors) %>% names()
               
for(i in colstomodify){
  sites_lr[,i][sites_lr[,i] < -5000] <- NA
}                         
```

### Investigate NA's
```{r}
locate.nas(sites_lr)
# There are some mystery NA's in here that need to be looked at.
# The columns that have all the NA's come from sites_aa_5m. Were these NA's in the original file, or were they the result of a faulty join?
locate.nas(sites_aa_5m) # There are no NA's in the original file, so there must have been a problem joining. 
# Immediately after joining, there are no NA's in the relevant columns. They must have been introduced later on.
# Turns out the NA's were properly introduced when I went through and changed 0's to NA's. Indeed, there were 947 sites in the original sites_aa_5m dataset that had 0 across all the columns, which means that the intersection with the layer failed. The NA's are a result of the original data, then, not a mistake on my end.
```


### Exclude UXO and TWZ
```{r}
sites_lr <- sites_lr %>% filter(!(stratum %in% c("UXO", "TWZ")))
```

### Remove pseudo-shoreline observations
```{r}
ps_lr <- sites_lr %>% filter(summary == "7")
ps_lr %$% table(stratum, pool)
sites_lr <- sites_lr %>% filter(summary != "7") # a code of 7 in the summary column means that the observations are pseudo-shoreline and should be excluded.
```

### Remove NA's for snag
```{r}
sites_lr <- sites_lr %>%
  filter(!is.na(snag))
```

## Save data
```{r}
#save(sites_lr, file = "../data/outputs/sites_lr.Rda")
```

# __________________
# Join data
## Align column names
### Check column names
```{r}
names(sites_ur)
names(sites_lr)
```

### Fix inconsistencies
```{r}
sites_lr <- sites_lr %>% dplyr::select(-c("OBJECTID_1", "RowID_"))
sites_lr <- sites_lr %>% mutate(river_mile = NA)
names(sites_lr) %in% names(sites_ur)
names(sites_ur) %in% names(sites_lr)
names(sites_ur) == names(sites_lr)
# Now they have the same names but in a different order
sites_lr <- sites_lr[,names(sites_ur)] # align names
```

## Add ID column
```{r}
sites_lr <- sites_lr %>% mutate(upperlower = "lower")
sites_ur <- sites_ur %>% mutate(upperlower = "upper")
```

## Remove ALT and OR1
```{r}
sites_lr <- sites_lr %>% 
  filter(!(pool %in% c("ALT", "OR1"))) %>%
  droplevels()
sites_ur <- sites_ur %>% 
  filter(!(pool %in% c("ALT", "OR1"))) %>%
  droplevels()
```

## Remove bad coords
### Upper
```{r}
sites_ur <- sites_ur %>%
  dplyr::select(-c("utm_e", "utm_n"))
# visualize zone
sites_ur %>%
  ggplot(aes(x = zone15e, y = zone15n, col = stratum))+
  geom_point()
# Remove bad points
sites_ur <- sites_ur %>%
  filter(zone15n > 2000000)
# visualize
sites_ur %>%
  ggplot(aes(x = zone15e, y = zone15n, col = stratum))+
  geom_point()
```
### Lower
```{r}
sites_lr <- sites_lr %>%
  dplyr::select(-c("utm_e", "utm_n"))
# visualize zone
sites_lr %>%
  ggplot(aes(x = zone15e, y = zone15n, col = stratum))+
  geom_point()
# Remove bad points
sites_lr <- sites_lr %>%
  filter(zone15e > 600000 & zone15e < 900000)
# visualize
sites_lr %>%
  ggplot(aes(x = zone15e, y = zone15n, col = stratum))+
  geom_point()
```

## Perform join
```{r}
sites_upperlower <- rbind(sites_lr, sites_ur)
sites_upperlower %>% filter(!(stratum %in% c("MCB-W", "BWC-O"))) %>%
  filter(!is.na(snag)) %>%
  locate.nas()
```

## NA pools
Why do we still have some pool NA's?
```{r}
napools <- sites_upperlower %>% filter(is.na(pool))
head(napools)
table(napools$upperlower, exclude = FALSE)
# all the NA pools are in the lower river, for some reason.
table(napools$stratum)
# make a map of the NA points
napools %>%
  ggplot(aes(x = zone15e, y = zone15n))+
  geom_point()
# map of the lower river for comparison
sites_lr %>% filter(!is.na(pool)) %>%
  ggplot(aes(x = zone15e, y = zone15n, col = pool))+
  geom_point()
# Assign pools based on zone15n
sites_upperlower <- sites_upperlower %>%
  mutate(fixedpool = case_when(is.na(pool) ~ TRUE,
                               TRUE ~ FALSE)) %>%
  mutate(pool = case_when(is.na(pool) & zone15n > 4400000 ~ "LAG",
                          is.na(pool) & zone15n < 4400000 & zone15n > 4200000 ~ "P26",
                          is.na(pool) & zone15n < 4200000 ~ "OR2",
                          TRUE ~ as.character(pool)))
```

# Save psedoshoreline obs
```{r}
ps_upperlower <- bind_rows(ps_ur, ps_lr)
#save(ps_upperlower, file = "../data/outputs/ps_upperlower.Rda")
```


## Save data
```{r}
#save(sites_upperlower, file = "../data/outputs/sites_upperlower.Rda")
#save(sites_ur, file = "../data/outputs/sites_ur.Rda")
#save(sites_lr, file = "../data/outputs/sites_lr.Rda")
# As CSV's
#write.csv(sites_upperlower, "../data/outputs/sites_upperlower.csv", row.names = F)
#write.csv(sites_ur, "../data/outputs/sites_ur.csv", row.names = F)
#write.csv(sites_lr, "../data/outputs/sites_lr.csv", row.names = F)
```