---
title: "descriptive_stats"
author: "Kaija Gahm"
date: "1/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Source libraries and functions
```{r}
source("../R_scripts/libraries.R")
source("../R_scripts/ownfunctions.R")
```

# Load data (combined upper/lower river)
```{r}
load("../data/outputs/sites_upperlower.Rda")
load("../data/outputs/sites_ur.Rda")
load("../data/outputs/sites_lr.Rda")
```

# Data Cleaning
## Remove the river mile column
```{r}
head(sites_upperlower)
sites_upperlower <- sites_upperlower %>% dplyr::select(-river_mile)
```

## Exclude strata: BWC-O and MCB-W
```{r}
sites_upperlower <- sites_upperlower %>% filter(stratum %notin% c("BWC-O", "MCB-W")) %>% droplevels()
```

## Exclude sites NA for snag and/or pool
```{r}
sites_upperlower <- sites_upperlower %>% filter(!is.na(snag), !is.na(pool))
```

# Summaries
## Overall
```{r}
summary_all <- sites_upperlower %>% dplyr::summarize(npoints = n(),
                                              wood = sum(snag == 1),
                                              nowood = sum(snag == 0),
                                              propwood = wood/npoints)
```

## By pool
```{r}
# Summary
summary_pool <- sites_upperlower %>% 
  group_by(pool) %>%
  dplyr::summarize(npoints = n(),
            wood = sum(snag == 1),
            nowood =  sum(snag == 0),
            propwood = wood/npoints)

# Test for significant differences
chisq.test(summary_pool[,c("wood", "nowood")])

# Post-hoc test with Bonferroni correction
pairwise.prop.test(x = summary_pool$wood, n = summary_pool$npoints, 
                   p.adjust.method = "bonferroni")
```

## By year and pool
```{r}
# Summary
summary_yearpool <- sites_upperlower %>% group_by(pool, year) %>%
  dplyr::summarize(npoints = n(),
            wood = sum(snag == 1),
            nowood =  sum(snag == 0),
            propwood = wood/npoints) %>% ungroup()

# Plot
summary_yearpool %>% ggplot(aes(x = year, y = propwood, col = pool, group = pool))+
  geom_line()+
  theme_minimal()
```

## By stratum
```{r}
# Summary
summary_stratum <- sites_upperlower %>% group_by(stratum) %>% 
  dplyr::summarize(npoints = n(),
            wood = sum(snag == 1),
            nowood = sum(snag == 0),
            propwood = wood/npoints)

# Calculate confidence intervals for these proportions
cis <- BinomCI(summary_stratum$wood, n = summary_stratum$npoints, conf.level = 0.95)
summary_stratum <- cbind(summary_stratum, cis[,2:3])
row.names(summary_stratum) <- NULL

# Test for significant differences
chisq.test(summary_stratum[,c("wood", "nowood")])

# Post-hoc test with Bonferroni correction
pairwise.prop.test(x = summary_pool$wood, n = summary_pool$npoints, 
                   p.adjust.method = "bonferroni")

# Plot of proportions and confidence intervals by stratum
summary_stratum %>% ggplot(aes(x = stratum, y = propwood))+
  geom_point(size = 3)+
  geom_errorbar(aes(ymin = lwr.ci,
                    ymax = upr.ci),
                width = 0,
                size = 1.2)+
  theme_minimal()+
  theme(text = element_text(size = 16))+
  ylab("Proportion of points with wood")+
  xlab("Habitat Stratum")+
  ggtitle("Wood proportion by stratum")

```

## By stratum and pool
```{r}
# Summary
summary_stratumpool <- sites_upperlower %>% group_by(pool, stratum) %>%
  dplyr::summarize(npoints = n(),
            wood = sum(snag == 1),
            nowood = sum(snag == 0),
            propwood = wood/npoints) %>% as.data.frame()

# Test for significant differences
chisq.test(summary_stratumpool[,c("wood", "nowood")])

# Confidence intervals
cis <- BinomCI(summary_stratumpool$wood, n = summary_stratumpool$npoints, conf.level = 0.95) %>% as.data.frame()
summary_stratumpool <- summary_stratumpool %>% mutate(lwr.ci = cis$lwr.ci,
                                                      upr.ci = cis$upr.ci)
row.names(summary_stratumpool) <- NULL

# Plot
summary_stratumpool %>% ggplot(aes(x = stratum, y = propwood, color = pool))+
  geom_point(size = 3,
             position = position_dodge(width = 0.3))+
  geom_errorbar(aes(ymin = lwr.ci, ymax = upr.ci),
                width = 0, size = 1.2, position = position_dodge(width = 0.3))+
  theme_minimal()+
  theme(text = element_text(size = 16))+
  ylab("Proportion of points with wood")+
  xlab("Habitat Stratum")

```

## By pool, year, and stratum
```{r}
# Summary
summary_poolyearstratum <- sites_upperlower %>% 
  group_by(pool, year, stratum) %>% 
  dplyr::summarize(npoints = n(),
            wood = sum(snag == 1),
            nowood = sum(snag == 0),
            propwood = wood/npoints) %>%
  as.data.frame()
head(summary_poolyearstratum)   

# Plot of proportions by stratum and pool
summary_poolyearstratum %>% ggplot(aes(x = year, y = propwood, group = stratum))+
  geom_line(aes(linetype = stratum))+
  ylim(0,1)+
  theme_minimal()+
  xlab("Year")+
  ylab("Proportion of points with wood")+
  facet_wrap(~pool)

# Plot of proportions by stratum and pool, only the two most common strata
summary_poolyearstratum %>% filter(stratum %in% c("MCB-U", "SCB")) %>%
  ggplot(aes(x = year, y = propwood, group = stratum))+
  geom_line(size = .65, aes(linetype = stratum))+
  theme_minimal()+
  xlab("Year")+
  ylab("Proportion of points with wood")+
  facet_wrap(~pool)

#Violin plots of annual variation by stratum and pool 
# All observations, by stratum and pool
summary_poolyearstratum %>% ggplot(aes(x = stratum, y = propwood))+
  geom_violin()+
  geom_boxplot(width = 0.1,
               fill = "black",
               outlier.color = NA)+
  stat_summary(fun.y = median,
               geom = "point",
               fill = "white",
               shape = 21,
               size = 2.5)+
  facet_wrap(~pool)

# Pools, including those with > 9 observations
summary_poolyearstratum %>% filter(npoints > 9) %>%
  ggplot(aes(x = pool, y = propwood))+
  geom_violin()+
  geom_boxplot(width = 0.1,
               fill = "black",
               outlier.color = NA)+
  stat_summary(fun.y = median,
               geom = "point",
               fill = "white",
               shape = 21,
               size = 2.5)

# Stratum and pool, greater than 10 observations
summary_poolyearstratum %>% filter(npoints > 9) %>%
  ggplot(aes(x = stratum, y = propwood))+
  geom_violin()+
  geom_boxplot(width=0.1,
               fill="black",
               outlier.colour=NA)+
  stat_summary(fun.y=median, 
               geom="point",
               fill="white",
               shape=21,
               size=2.5)+
  facet_wrap(~pool)
```

# Time series
## Set plotting parameters
```{r}
ylims <- c(0,1)
```

## Overall
### Examine by pool
```{r}
summary_yearpool %>% ggplot(aes(x = as.numeric(year), y = propwood))+
  geom_point()+
  stat_smooth(method = "loess", col = "red", se = F)+
  theme(text = element_text(size = 20))+
  facet_wrap(~pool)
```

### Pool 4
#### All years
```{r}
p4 <- summary_yearpool %>% filter(pool == "P04") %>% dplyr::select(year, propwood) %>% mutate(year = as.numeric(year))
plot(propwood ~ year, data = p4, pch = 19)
x <- p4$year
y <- p4$propwood
lin.mod <- lm(y ~ x)
seg.mod <- segmented(lin.mod, seg.Z = ~x, data = p4, psi = 1999) # singular

plot(y ~ x, 
     pch = 19, 
     cex = 1.5, 
     xlab = "Year",
     ylab = "Proportion with LW",
     main = "Pool 4")
plot(seg.mod, add = T)
```

#### After 1999
```{r}
p4b <- p4 %>% filter(year >= 1999)
x <- p4b$year
y <- p4b$propwood
lin.mod <- lm(y ~ x)
plot(y ~ x, pch = 19,
     cex = 1.5,
     xlab = "Year",
     ylab = "Proportion with LW",
     main = "Pool 4")
abline(lin.mod)

p4b %>% ggplot(aes(x = year, y = propwood))+
  geom_point()+
  stat_smooth(method = "lm", col = "red")+
  theme_minimal()
```

#### Tests
```{r}
pettitt.test(p4$propwood)  #first 6 points is separate
mk.test(p4$propwood) #monotonic trend? 
```
#### Mean before change
```{r}
plot(propwood ~ year, data = p4,
     pch = 19,
     cex = 1.5,
     ylim = ylims,
     xlab = "Year",
     ylab = "Proportion with LW",
     main = "Pool 4",
     cex.lab = 1.5,
     cex.axis = 1.25)
abline(v = 1998.5, col = "red", lwd = 3, lty = 2)

lm1 <- lm(propwood ~ year, data = p4[1:6,])
x0 = 1992.5
y0 = (x0 * lm1$coefficients[[2]]) + lm1$coefficients[[1]]
x1 = 1998.5
y1 = (x1 * lm1$coefficients[[2]] + lm1$coefficients [[1]])
         #segments(x0,y0,x1,y1,lty=1,lwd=3,col="darkgrey")
summary(lm1)
mn <- p4[1:6,] %>% pull(propwood) %>% mean()

segments(x0, mn, x1, mn, lwd = 3, lty = 2, col = "darkgrey")
text(1995, 0.25, labels = "mean = 0.487", cex = 1.5) # 2/15/20 THIS IS NO LONGER CORRECT
```






## Aquatic areas by pool
### P4
#### Exploratory
```{r}
p4 <- summary_poolyearstratum %>% filter(pool == "P04") %>% dplyr::select(year, stratum, propwood)

  # exploratory first...
  p4 %>% ggplot(aes(x = as.numeric(year), y = propwood))+
    geom_point()+
    stat_smooth(method = "loess", col = "red")+
    theme_minimal()+
    theme(text = element_text(size = 20))+
    facet_wrap(~stratum)
  
  # examine habitats separately
  p4.bwc <- p4 %>% filter(stratum == "BWC-S")
  pettitt.test(p4.bwc$propwood) # first 6 points is separate [edit 2/15/20: i don't know how to interpret this test, just reproducing the comments as they were before.]
  mk.test(p4.bwc$propwood) # monotonic trend?
  
  p4.mcb <- p4 %>% filter(stratum == "MCB-U")
  pettitt.test(p4.mcb$propwood) # first 6 points is separate 
  mk.test(p4.mcb$propwood) # monotonic trend?

  p4.scb <- p4 %>% filter(stratum == "SCB")
  pettitt.test(p4.scb$propwood) # first 6 points is separate 
  mk.test(p4.scb$propwood) # monotonic trend?
```

#### Plots for export
```{r}
#Plot for export
ah.pch <- c(NA, 3, NA, NA, NA, 19, NA, 1)[as.numeric(p4$stratum)]
plot(propwood ~ year, 
     data = p4, 
     cex = 1.75,
     xlab = "Year",
     ylab = "Proportion with LW",
     main = "Pool 4",
     cex.lab = 1.5,
     cex.axis = 1.25,
     ylim = ylims,
     pch = ah.pch)
abline(v = 1998.5, col = "black", lty = 1, lwd = 1) #MCB - or lwd = 3?
abline(v = 1998.5, col = "grey", lty = 2, lwd = 1) #BWC
abline(v = 2000.5, col = "grey", lty = 1, lwd = 1) # SCB
points(propwood ~ year, data = p4, cex = 1.75, pch = ah.pch) # Kaija 2/15/20: why is this necessary if we already specified the points and point types in the original plot call?
```





