---
title: "descriptive_stats"
author: "Kaija Gahm"
date: "1/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Source libraries and functions
```{r echo = F, message = F}
source("../R_scripts/libraries.R")
source("../R_scripts/ownfunctions.R")
```

# Load data (combined upper/lower river)
```{r}
load("../data/outputs/sites_upperlower.Rda")
load("../data/outputs/sites_ur.Rda")
load("../data/outputs/sites_lr.Rda")
```

# Data Cleaning
## Examine strata
```{r}
sites_upperlower <- sites_upperlower %>% droplevels()
sites_upperlower %$% table(stratum)
```

## Check for NA's in snag or pool
```{r}
sites_upperlower %>% filter(is.na(snag) | is.na(pool)) # good!
```

<!-- # Update strata -->
<!-- ## Exclude MCB-W and BWC-O -->
<!-- ```{r} -->
<!-- sites_upperlower <- sites_upperlower %>%  -->
<!--   filter(!(stratum %in% c("MCB-W", "BWC-O"))) %>% -->
<!--   droplevels() -->

<!-- sites_upperlower <- sites_upperlower %>% -->
<!--   mutate(stratum = case_when(stratum == "BWC-S" ~ "BWC", -->
<!--                              stratum == "IMP-S" ~ "IMP", -->
<!--                              stratum == "MCB-U" ~ "MCB", -->
<!--                              TRUE ~ as.character(stratum))) -->

<!-- table(sites_upperlower$stratum, exclude = NULL) -->
<!-- ``` -->


# Investigate numbers
The numbers above are just slightly _higher_ than the ones in the manuscript. Part of that can probably be accounted for by the missing pool numbers, which I just added, but some problems are left unexplained.

Load old data to verify it against the current data
```{r}
load("~/Downloads/all_reduced_clean.Rda")
load("~/Downloads/all2_reduced_clean.Rda")

# rename data
old_upper <- all_reduced_clean
old_lower <- all2_reduced_clean

# created combined dataset (old data)
old_upperlower <- bind_rows(old_upper, old_lower)

# remove some environmental vars
old_upperlower <- old_upperlower %>%
  dplyr::select(-c(near_terr_dist, near_terr_class_7, near_terr_name, near_forest_dist))
names(old_upperlower)

# Reassign strata
old_upperlower <- old_upperlower %>%
  mutate(stratum = case_when(stratum == "BWC-S" ~ "BWC",
                             stratum == "IMP-S" ~ "IMP",
                             stratum == "MCB-U" ~ "MCB",
                             TRUE ~ as.character(stratum))) %>%
  filter(stratum != "BWC-O")

# get stratum breakdown
table(old_upperlower$stratum, exclude = NULL) #these are the numbers we originally had in the manuscript

# get stratum breakdown for current sites
table(sites_upperlower$stratum, exclude = NULL)

# comparing these two tables shows that there are discrepancies in almost every pool-by-stratum category. So the issue is not confined to a few. Thus probably a more general data quality issue, or having to do with a stratum- or pool-nonspecific factor. 
 
# differences between new and old sites, broken down by stratum and pool
with(sites_upperlower, table(stratum, pool)) %>% as.matrix() - with(old_upperlower, table(stratum, pool)) %>% as.matrix()

# Now let's look only at the ones in the new dataset that aren't the same as the ones in the old
# by barcode
sites_upperlower %>%
  filter(!(barcode %in% old_upperlower$barcode)) %$%
  table(stratum, pool)

old_upperlower %>%
  filter(!(barcode %in% sites_upperlower$barcode)) %$%
  table(stratum, pool)

# It's concerning that this goes both ways. There are some new sites that weren't represented in the old sites, and some old sites that weren't represented in the new sites
# Hopefully, the old sites not represented in the new sites will be the pseudo-shoreline sites. 

# investigate the old sites not represented in the new sites
old_df <- old_upperlower %>%
  filter(!(barcode %in% sites_upperlower$barcode)) %$%
  table(stratum, pool) %>%
  as.data.frame()

# got these data manually from going back and examining the pseudoshoreline observations before I removed them.
ps_df <- data.frame(stratum = c("BWC", "MCB", "SCB", "IMP", "MCB", "IMP", "BWC", "MCB", "SCB", "MCB", "SCB", "BWC", "IMP", "MCB", "SCB"),
                    pool = c("P04", "P04", "P04", "P08", "P08", "P13", "LAG", "LAG", "LAG", "OR2", "OR2", "P26", "P26", "P26", "P26"),
                    freq_ps = c(208, 22, 40, 5, 1, 1, 117, 81, 88, 3, 1, 9, 1, 1, 4))

diffs_df <- left_join(old_df, ps_df, by = c("stratum", "pool"))
diffs_df$freq_ps[is.na(diffs_df$freq_ps)] <- 0 # change NA to 0

diffs_df <- diffs_df %>%
  mutate(leftover_problems = Freq-freq_ps)

diffs_df # okay, looking at the leftover_problems column, we have a couple negative numbers. That is weird. Means that some of what showed up as pseudo-shoreline observations in the new data got removed by other means from the old data (back when we weren't yet considering pseudo-shoreline.)
```

# Investigation of obs. present in old, missing from new
```{r}
# Load pseudoshoreline data
load("../data/outputs/ps_upperlower.Rda")

# get missing data each way
missing_from_new <- old_upperlower %>% filter(!(barcode %in% sites_upperlower$barcode))

missing_from_new$barcode %in% ps_upperlower$barcode # most of the sites missing from the new data are attributable to the pseudoshoreline obs

missing_from_new <- missing_from_new %>%
  mutate(ps = case_when(barcode %in% ps_upperlower$barcode ~ T,
                        TRUE ~ FALSE))
troublesome_barcodes <- missing_from_new %>% filter(ps == FALSE) %>% pull(barcode) # pull barcodes from the ones that remain problematic

# plot the new sites, colored by pool. Add the still-problematic sites, to see if this is a coordinates issue. 
sites_upperlower %>%
  ggplot(aes(x = zone15e, y = zone15n, col = pool))+
  geom_point()+
  geom_point(data = missing_from_new %>% filter(!ps), aes(x = utm_e, y = utm_n), col = "black") # the still missing sites fall within reasonable bounds for coordinates, but interestingly they're all in OR2. That's interesting. 

troublesome_barcodes # here are the barcodes of the ones that are still problematic, after accounting for the pseudoshoreline obs.

missing_from_new %>% filter(barcode %in% troublesome_barcodes)
# I went back through the data cleaning pipeline to figure out when these four observations disappeared. Seems that in the new data, they were classified as OR1, not OR2, and so they were removed in the step where I excluded ALT and OR1.

# Next step is to go back through the data pipeline and see if they were classified as OR1 in the original data, or if they somehow got changed from OR2 to OR1. 
# Ok, in the original fish data, these were classified as OR, not OR1 or OR2. When the aa dataset got joined in, they were given the classification OR1, and were thus excluded.
# Lookin at missing_from_new now (which is derived from the old data), we see that despite "pool" being listed as OR2, the uniq_id shows these observations as belonging to OR1. Thus, I'm going to conclude that the OR2 in the old data was a mistake, and that these observations are supposed to be part of OR1. Therefore, it was correct to exclude them from the new data, so nothing is wrong here.

# I have now investigated all observations that were present in the old data but are missing from the new data. Now time to do the other side: observations present in the new data that were missing from the old data.
```

# Investigation of obs. present in new, missing from old
```{r}
missing_from_old <- sites_upperlower %>% filter(!(barcode %in% old_upperlower$barcode))
missing_from_old # looks like all these sites are NA for a bunch of environmental predictors. Probably I filtered these out in the old observations, but left them in in the new.

# Another thing I notice here is that the observations where fixedpool == FALSE (i.e. which already had pools; I didn't have to assign them based on utm coordinates) have less missing data than the ones where fixedpool == TRUE. That makes me think that the join to the aquatic areas dataset must have been based on pools. 

# Going to go back and try assigning the missing pools based on utm coordinates earlier in the process, before the join, to see if we can get some of that information back. 

# went back and checked: the join was in fact not based on pool. And I've noted that in some places, even when fixedpool == FALSE, the information is still missing. So I think these stem from a common cause of the intersection being bad.

# also went back into the old data cleaning code and saw that in fact, we did want to exclude the ones that are NA for a bunch of these environmental variables (didn't intersect the aquatic areas layer). I'll go back and do that now. 

# excluded observations that were NA, aka didn't intersect with the aquatic areas.

# After having done that, there are still a bunch that are missing from old but included in new. Have to figure out why this is.

sites_upperlower %$% table(stratum) # we actually now have lower numbers than we did in the old data, which is good, given the exclusion of pseudo-shoreline sites. But then why is there still a discrepancy?

missing_from_old %$% table(stratum, pool) # most of these are from BWC in LAG. Is it possible that we just had some more LAG observations added to the files or something?
```

In any case, for now I'm going to go forward with the data as it is. Can address these concerns in the meeting.

# Summaries
## Overall
```{r}
summary_all <- sites_upperlower %>% dplyr::summarize(npoints = n(),
                                              wood = sum(snag == 1),
                                              nowood = sum(snag == 0),
                                              propwood = wood/npoints)

summary_all
```

## By pool
```{r}
# Summary
summary_pool <- sites_upperlower %>% 
  group_by(pool) %>%
  dplyr::summarize(npoints = n(),
            wood = sum(snag == 1),
            nowood =  sum(snag == 0),
            propwood = wood/npoints)

# Test for significant differences
chisq.test(summary_pool[,c("wood", "nowood")])

# Post-hoc test with Bonferroni correction
pairwise.prop.test(x = summary_pool$wood, n = summary_pool$npoints, 
                   p.adjust.method = "bonferroni")
```

## By year and pool
```{r}
# Summary
summary_yearpool <- sites_upperlower %>% group_by(pool, year) %>%
  dplyr::summarize(npoints = n(),
            wood = sum(snag == 1),
            nowood =  sum(snag == 0),
            propwood = wood/npoints) %>% ungroup()

# Plot
summary_yearpool %>% ggplot(aes(x = year, y = propwood, col = pool, group = pool))+
  geom_line()+
  theme_minimal()
```

## By stratum
```{r}
# Summary
summary_stratum <- sites_upperlower %>% group_by(stratum) %>% 
  dplyr::summarize(npoints = n(),
            wood = sum(snag == 1),
            nowood = sum(snag == 0),
            propwood = wood/npoints)

# Calculate confidence intervals for these proportions
cis <- BinomCI(summary_stratum$wood, n = summary_stratum$npoints, conf.level = 0.95)
summary_stratum <- cbind(summary_stratum, cis[,2:3])
row.names(summary_stratum) <- NULL

# Test for significant differences
chisq.test(summary_stratum[,c("wood", "nowood")])

# Post-hoc test with Bonferroni correction
pairwise.prop.test(x = summary_pool$wood, n = summary_pool$npoints, 
                   p.adjust.method = "bonferroni")

# Plot of proportions and confidence intervals by stratum
summary_stratum %>% ggplot(aes(x = stratum, y = propwood))+
  geom_point(size = 3)+
  geom_errorbar(aes(ymin = lwr.ci,
                    ymax = upr.ci),
                width = 0,
                size = 1.2)+
  theme_minimal()+
  theme(text = element_text(size = 16))+
  ylab("Proportion of points with wood")+
  xlab("Habitat Stratum")+
  ggtitle("Wood proportion by stratum")

summary_stratum
```

## By stratum and pool
```{r}
# Summary
summary_stratumpool <- sites_upperlower %>% group_by(pool, stratum) %>%
  dplyr::summarize(npoints = n(),
            wood = sum(snag == 1),
            nowood = sum(snag == 0),
            propwood = wood/npoints) %>% as.data.frame()
summary_stratumpool

# Test for significant differences
chisq.test(summary_stratumpool[,c("wood", "nowood")])

# Confidence intervals
cis <- BinomCI(summary_stratumpool$wood, n = summary_stratumpool$npoints, conf.level = 0.95) %>% as.data.frame()
summary_stratumpool <- summary_stratumpool %>% mutate(lwr.ci = cis$lwr.ci,
                                                      upr.ci = cis$upr.ci)
row.names(summary_stratumpool) <- NULL

# Plot
summary_stratumpool %>% ggplot(aes(x = stratum, y = propwood, color = pool))+
  geom_point(size = 3,
             position = position_dodge(width = 0.3))+
  geom_errorbar(aes(ymin = lwr.ci, ymax = upr.ci),
                width = 0, size = 1.2, position = position_dodge(width = 0.3))+
  theme_minimal()+
  theme(text = element_text(size = 16))+
  ylab("Proportion of points with wood")+
  xlab("Habitat Stratum")

nrow(ps_upperlower)
```

## By pool, year, and stratum
```{r}
# Summary
summary_poolyearstratum <- sites_upperlower %>% 
  group_by(pool, year, stratum) %>% 
  dplyr::summarize(npoints = n(),
            wood = sum(snag == 1),
            nowood = sum(snag == 0),
            propwood = wood/npoints) %>%
  as.data.frame()
head(summary_poolyearstratum)   

# Plot of proportions by stratum and pool
summary_poolyearstratum %>% ggplot(aes(x = year, y = propwood, group = stratum))+
  geom_line(aes(linetype = stratum))+
  ylim(0,1)+
  theme_minimal()+
  xlab("Year")+
  ylab("Proportion of points with wood")+
  facet_wrap(~pool)

# Plot of proportions by stratum and pool, only the two most common strata
summary_poolyearstratum %>% filter(stratum %in% c("MCB-U", "SCB")) %>%
  ggplot(aes(x = year, y = propwood, group = stratum))+
  geom_line(size = .65, aes(linetype = stratum))+
  theme_minimal()+
  xlab("Year")+
  ylab("Proportion of points with wood")+
  facet_wrap(~pool)

#Violin plots of annual variation by stratum and pool 
# All observations, by stratum and pool
summary_poolyearstratum %>% ggplot(aes(x = stratum, y = propwood))+
  geom_violin()+
  geom_boxplot(width = 0.1,
               fill = "black",
               outlier.color = NA)+
  stat_summary(fun.y = median,
               geom = "point",
               fill = "white",
               shape = 21,
               size = 2.5)+
  facet_wrap(~pool)

# Pools, including those with > 9 observations
summary_poolyearstratum %>% filter(npoints > 9) %>%
  ggplot(aes(x = pool, y = propwood))+
  geom_violin()+
  geom_boxplot(width = 0.1,
               fill = "black",
               outlier.color = NA)+
  stat_summary(fun.y = median,
               geom = "point",
               fill = "white",
               shape = 21,
               size = 2.5)

# Stratum and pool, greater than 10 observations
summary_poolyearstratum %>% filter(npoints > 9) %>%
  ggplot(aes(x = stratum, y = propwood))+
  geom_violin()+
  geom_boxplot(width=0.1,
               fill="black",
               outlier.colour=NA)+
  stat_summary(fun.y=median, 
               geom="point",
               fill="white",
               shape=21,
               size=2.5)+
  facet_wrap(~pool)

table(summary_poolyearstratum$stratum)
summary_poolyearstratum %>% filter(pool == "P04", stratum == "BWC-S") %>% 
  ggplot(aes(x = year, y = propwood))+
  geom_point()
```

### Save summaries
```{r}
save(summary_poolyearstratum, file = "../data/outputs/summary_poolyearstratum.Rda")
save(summary_all, file = "../data/outputs/summary_all.Rda")
save(summary_pool, file = "../data/outputs/summary_pool.Rda")
save(summary_yearpool, file = "../data/outputs/summary_yearpool.Rda")
save(summary_stratum, file = "../data/outputs/summary_stratum.Rda")
save(summary_stratumpool, file = "../data/outputs/summary_stratumpool.Rda")
```


# Time series
## Set plotting parameters
```{r}
ylims <- c(0,1)
```

## Function for making models
```{r}
makemod <- function(model, x){
  y0 <- (x * model$coefficients[[2]]) + model$coefficients[[1]]
  return(y0)
}
```

## Overall
### Examine by pool
```{r}
summary_yearpool %>% ggplot(aes(x = as.numeric(year), y = propwood))+
  geom_point()+
  stat_smooth(method = "loess", col = "red", se = F)+
  facet_wrap(~pool)+
  xlab("Year")+
  ylab("Proportion with LW")
```

### Pool 4
#### All years
```{r}
p4 <- summary_yearpool %>% filter(pool == "P04") %>% dplyr::select(year, propwood) %>% mutate(year = as.numeric(year))
p4 %>% ggplot(aes(x = year, y = propwood))+
  geom_point()

lin.mod <- lm(propwood ~ year, data = p4)
seg.mod <- segmented(lin.mod, data = p4, psi = 1999) # singular
p4$seg_fit <- broken.line(seg.mod)$fit # add fitted values to the data frame.

baseplot_p4 <- p4 %>% ggplot(aes(x = year, y = propwood))+
  geom_point()+
  ylim(0, 1)+
  xlab("Year")+
  ylab("Proportion with LW")+
  ggtitle("Pool 4")+
  theme_minimal()

baseplot_p4 + geom_line(aes(y = seg_fit), lwd = 0.3)


plot(propwood ~ year, data = p4, # compare this to the above for how they connect the lines... weird that it's different.
     pch = 19, 
     cex = 1.5, 
     xlab = "Year",
     ylab = "Proportion with LW",
     main = "Pool 4")
plot(seg.mod, add = T)
```

#### After 1999
```{r}
p4b <- p4 %>% filter(year >= 1999)

p4b %>% ggplot(aes(x = year, y = propwood))+
  geom_point()+
  stat_smooth(method = "lm", col = "red", lwd = 0.3, se = F)+
  theme_minimal()
```

#### Tests
```{r}
pt <- pettitt.test(p4$propwood)  
str(pt)
(changepoint <- pt$estimate) # First 8 points separate
changeyear <- mean(c(p4$year[changepoint], p4$year[changepoint+1])) # get year for change in means

mk.test(p4$propwood) # Monotonic trend? [INTERPRETATION?]
```

#### Mean before change
```{r}
baseplot_p4 + geom_vline(aes(xintercept = changeyear), 
                         col = "red",
                         lty = 2)

# I don't really know what's going on with these models... can we leave them out?
lm1 <- lm(propwood ~ year, data = p4[1:changepoint,])
begyear <- 1992.5 # where does this come from? Prior knowledge?
y0 <- makemod(lm1, begyear)
y1 <- makemod(lm1, changeyear)
summary(lm1)

lm2 <- lm(propwood ~ year, data = p4[(changepoint + 1):nrow(p4),])
y0 <- makemod(lm2, changeyear)
endyear <- max(p4$year) + 0.5
y1 <- makemod(lm2, endyear)

mn1 <- p4[1:changepoint,] %>% pull(propwood) %>% mean()
mn2 <- p4[(changepoint + 1):nrow(p4),] %>% pull(propwood) %>% mean()

baseplot_p4 + geom_vline(aes(xintercept = changeyear), 
                         col = "red",
                         lty = 2) + 
  geom_segment(aes(x = begyear, y = mn1, xend = changeyear, yend = mn1),
               lty = 2,
               col = "darkgrey")+
  geom_segment(aes(x = changeyear, y = mn2, xend = endyear, yend = mn2),
               lty = 2,
               col = "darkgrey")+
  geom_text(aes(x = 1995, y = mn1 - 0.05, label = paste0("mean = ", round(mn1, 2))), size = 4,
            col = "darkgrey",
            check_overlap = T)+
  geom_text(aes(x = 2006, y = mn2 - 0.05, label = paste0("mean = ", round(mn2, 2))), size = 4,
            col = "darkgrey",
            check_overlap = T)
```

### Pool 8
#### All years
```{r}
p8 <- summary_yearpool %>% filter(pool == "P08") %>% dplyr::select(year, propwood) %>% mutate(year = as.numeric(year))

lin.mod <- lm(propwood ~ year, data = p8)
seg.mod <- segmented(lin.mod, data = p8, seg.Z = ~ year, psi = 2003)

pt <- pettitt.test(p8$propwood) # first 6 points is separate
(changepoint <- pt$estimate) # First 8 points separate
changeyear <- mean(c(p8$year[changepoint], p8$year[changepoint + 1])) # get year for change in means
mk.test(p8$propwood) # monotonic trend? yes, highly significant trend

p8 %>% ggplot(aes(x = year, y = propwood))+
  geom_point()+
  xlab("Year")+
  ylab("Proportion with LW")+
  ggtitle("Pool 8")
```


```{r}
lm1 <- lm(propwood ~ year, data = p8[1:changepoint,])
lm2 <- lm(propwood ~ year, data = p8[(changepoint + 1):nrow(p8),])
slope1 <- lm1$coefficients[2] # extract slope
slope2 <- lm2$coefficients[2] # extract slope

p8 %>% ggplot(aes(x = year, y = propwood))+
  geom_point()+
  geom_vline(aes(xintercept = changeyear), 
                         col = "red",
                         lty = 2)+
  geom_smooth(data = p8[1:changepoint,], 
              method = "lm", 
              se = F, 
              col = "darkgrey",
              lwd = 0.5)+
  geom_smooth(data = p8[(changepoint + 1):nrow(p8),],
              method = "lm",
              se = F,
              col = "darkgrey",
              lwd = 0.5)+
  geom_text(aes(x = 1997, y = 0.7, label = paste0("slope = ", round(slope1, 3))), size = 4,
            col = "darkgrey",
            check_overlap = T)+
  geom_text(aes(x = 2010, y = 0.45, label = paste0("slope = ", round(slope2, 3))), size = 4,
            col = "darkgrey",
            check_overlap = T)
```

### Pool 13
#### All years
```{r}
p13 <- summary_yearpool %>% filter(pool == "P13") %>% dplyr::select(year, propwood) %>% mutate(year = as.numeric(year))

lin.mod <- lm(propwood ~ year, data = p13)
seg.mod <- segmented(lin.mod, data = p13, psi = 2005) # what is the psi??

pt <- pettitt.test(p13$propwood) # first 16 points is separate
(changepoint <- pt$estimate) # First 16 points separate
changeyear <- mean(c(p13$year[changepoint], p13$year[changepoint + 1])) # get year for change in means
mk.test(p13$propwood) # monotonic trend? no, not really.

p13 %>% ggplot(aes(x = year, y = propwood))+
  geom_point()+
  xlab("Year")+
  ylab("Proportion with LW")+
  ggtitle("Pool 13")
```
```{r}
mn1 <- p13 %>% filter(year < changeyear) %>% pull(propwood) %>% mean()
mn2 <- p13 %>% filter(year > changeyear) %>% pull(propwood) %>% mean()
begyear <- min(p13$year) - 0.5
endyear <- max(p13$year) + 0.5

p13 %>% ggplot(aes(x = year, y = propwood))+
  geom_point()+
  xlab("Year")+
  ylab("Proportion with LW")+
  ggtitle("Pool 13") + 
  geom_vline(aes(xintercept = changeyear), 
                         col = "red",
                         lty = 2) + 
  geom_segment(aes(x = begyear, y = mn1, xend = changeyear, yend = mn1),
               lty = 2,
               col = "darkgrey") +
  geom_segment(aes(x = changeyear, y = mn2, xend = endyear, yend = mn2),
               lty = 2,
               col = "darkgrey") +
  geom_text(aes(x = 1995, y = mn1 - 0.05, label = paste0("mean = ", round(mn1, 2))), size = 4,
            col = "darkgrey",
            check_overlap = T) +
  geom_text(aes(x = 2013, y = mn2 - 0.05, label = paste0("mean = ", round(mn2, 2))), size = 4,
            col = "darkgrey",
            check_overlap = T)
```

### Pool 26
#### All years
```{r}
p26 <- summary_yearpool %>% filter(pool == "P26") %>% dplyr::select(year, propwood) %>% mutate(year = as.numeric(year))

lin.mod <- lm(propwood ~ year, data = p26)
seg.mod <- segmented(lin.mod, data = p26, psi = 2000) # what is psi

pt <- pettitt.test(p26$propwood) # first 6 points is separate
(changepoint <- pt$estimate) # First 8 points separate
changeyear <- mean(c(p26$year[changepoint], p26$year[changepoint + 1])) # get year for change in means
mk.test(p26$propwood) # no trend

p26 %>% ggplot(aes(x = year, y = propwood))+
  geom_point()+
  xlab("Year")+
  ylab("Proportion with LW")+
  ggtitle("Pool 26")


x <- p26$year
y <- p26$propwood
plot(y ~ x, pch = 19,
     cex = 1.5,
     xlab = "Year",
     ylab = "Proportion with LW",
     main = "Pool 26")
plot(seg.mod, add = T)
summary(seg.mod)
summary(lin.mod)
slope(seg.mod)

pettitt.test(p26$propwood) # break in mean? NO
mk.test(p26$propwood) # monotonic trend? NO
(mean1 <- mean(p26$propwood)) #No trends/breaks whatsoever
     
plot(propwood ~ year, data = p26,
     pch = 19,
     cex = 1.75,
     xlab = "Year",
     ylab = "Proportion with LW",
     main = "Pool 26",
     cex.lab = 1.5,
     cex.axis = 1.25,
     ylim = ylims)
abline(h = mean(p26$propwood), col = "darkgrey", lwd = 2, lty = 2)
points(propwood ~ year, data = p26, pch = 19, cex = 1.75)
text(2015, 0.9, labels = "mean = 0.594", cex = 1.5)   
```








## Aquatic areas by pool
### P4
#### Exploratory
```{r}
p4 <- summary_poolyearstratum %>% filter(pool == "P04") %>% dplyr::select(year, stratum, propwood)

  # exploratory first...
  p4 %>% ggplot(aes(x = as.numeric(year), y = propwood))+
    geom_point()+
    stat_smooth(method = "loess", col = "red")+
    theme_minimal()+
    theme(text = element_text(size = 15))+
    facet_wrap(~stratum)+
    xlab("Year")+
    ylab("Proportion with LW")
  
  # examine habitats separately
  p4.bwc <- p4 %>% filter(stratum == "BWC-S")
  pettitt.test(p4.bwc$propwood) # first 6 points is separate [edit 2/15/20: i don't know how to interpret this test, just reproducing the comments as they were before.]
  mk.test(p4.bwc$propwood) # monotonic trend?
  
  p4.mcb <- p4 %>% filter(stratum == "MCB-U")
  pettitt.test(p4.mcb$propwood) # first 6 points is separate 
  mk.test(p4.mcb$propwood) # monotonic trend?

  p4.scb <- p4 %>% filter(stratum == "SCB")
  pettitt.test(p4.scb$propwood) # first 6 points is separate 
  mk.test(p4.scb$propwood) # monotonic trend?
```

#### Plots for export
```{r}
#Plot for export
ah.pch <- c(NA, 3, NA, NA, NA, 19, NA, 1)[as.numeric(p4$stratum)]
plot(propwood ~ year, 
     data = p4, 
     cex = 1.75,
     xlab = "Year",
     ylab = "Proportion with LW",
     main = "Pool 4",
     cex.lab = 1.5,
     cex.axis = 1.25,
     ylim = ylims,
     pch = ah.pch)
abline(v = 1998.5, col = "black", lty = 1, lwd = 1) #MCB - or lwd = 3?
abline(v = 1998.5, col = "grey", lty = 2, lwd = 1) #BWC
abline(v = 2000.5, col = "grey", lty = 1, lwd = 1) # SCB
points(propwood ~ year, data = p4, cex = 1.75, pch = ah.pch) # Kaija 2/15/20: why is this necessary if we already specified the points and point types in the original plot call?
```





