---
title: "descriptive_stats_2"
author: "Kaija Gahm"
date: "5/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Source libraries and functions
```{r echo = F}
source("../R_scripts/libraries.R")
source("../R_scripts/ownfunctions.R")
```

# Load data (combined upper/lower river)
```{r load-data}
load("../data/outputs/sites_upperlower.Rda")
load("../data/outputs/sites_ur.Rda")
load("../data/outputs/sites_lr.Rda")
```

# Data Cleaning
## Examine strata
```{r table-strata}
sites_upperlower <- sites_upperlower %>% droplevels()
nrow(sites_upperlower)
```

## Check for NA's in snag or pool
```{r check-nas}
sites_upperlower %>% filter(is.na(snag) | is.na(pool)) # good!
```

## Compare to old data
### Load old data
```{r load-old}
load("../data/outputs/all_reduced_clean.Rda")
load("../data/outputs/all2_reduced_clean.Rda")

# rename data
old_upper <- all_reduced_clean
old_lower <- all2_reduced_clean

# created combined dataset (old data)
old_upperlower <- bind_rows(old_upper, old_lower)
```

```{r investigate-upper}
nrow(sites_ur)
nrow(old_upper)

length(unique(sites_ur$barcode)) # there seems to be a unique barcode for each row
length(unique(old_upper$barcode))
```
Okay, we have differences on both sides: sites appearing in the new data that didn't appear in the old, and sites appearing in the old data that don't appear in the new.

Since the new data cleaning pipeline is pretty much as simple as I can get it, and given that there are many possible sources of error (pseudo-shoreline obs, and non-intersections with aquatic areas, to name two), I'm going to go with the data I have now and not investigate further.

# Summaries
## Overall
```{r}
summary_all <- sites_upperlower %>% dplyr::summarize(npoints = n(),
                                              wood = sum(snag == 1),
                                              nowood = sum(snag == 0),
                                              propwood = wood/npoints)

summary_all
```

## By pool
```{r}
# Summary
summary_pool <- sites_upperlower %>% 
  group_by(pool) %>%
  dplyr::summarize(npoints = n(),
            wood = sum(snag == 1),
            nowood =  sum(snag == 0),
            propwood = wood/npoints)

# Test for significant differences
chisq.test(summary_pool[,c("wood", "nowood")])

# Post-hoc test with Bonferroni correction
summary_pool <- summary_pool %>%
  mutate(pool = factor(pool, levels = c("4", "8", "13", "26", "LG", "OR"))) %>%
  arrange(pool)
pairwise.prop.test(x = summary_pool$wood, n = summary_pool$npoints, 
                   p.adjust.method = "bonferroni")

unique(summary_pool$pool)
```

## By year and pool
```{r}
# Summary
summary_yearpool <- sites_upperlower %>% group_by(pool, year) %>%
  dplyr::summarize(npoints = n(),
            wood = sum(snag == 1),
            nowood =  sum(snag == 0),
            propwood = wood/npoints) %>% ungroup()

# Plot
summary_yearpool %>% ggplot(aes(x = year, y = propwood, col = pool, group = pool))+
  geom_line()+
  theme_minimal()
```

## By stratum
```{r}
# Summary
summary_stratum <- sites_upperlower %>% group_by(stratum) %>% 
  dplyr::summarize(npoints = n(),
            wood = sum(snag == 1),
            nowood = sum(snag == 0),
            propwood = wood/npoints)

# Calculate confidence intervals for these proportions
cis <- BinomCI(summary_stratum$wood, n = summary_stratum$npoints, conf.level = 0.95)
summary_stratum <- cbind(summary_stratum, cis[,2:3])
row.names(summary_stratum) <- NULL

# Test for significant differences
chisq.test(summary_stratum[,c("wood", "nowood")])

# Post-hoc test with Bonferroni correction
pairwise.prop.test(x = summary_stratum$wood, n = summary_stratum$npoints, 
                   p.adjust.method = "bonferroni")
unique(summary_stratum$stratum)

# Plot of proportions and confidence intervals by stratum
summary_stratum %>% ggplot(aes(x = stratum, y = propwood))+
  geom_point(size = 3)+
  geom_errorbar(aes(ymin = lwr.ci,
                    ymax = upr.ci),
                width = 0,
                size = 1.2)+
  theme_minimal()+
  theme(text = element_text(size = 16))+
  ylab("Proportion of points with wood")+
  xlab("Habitat Stratum")+
  ggtitle("Wood proportion by stratum")

summary_stratum
```

## By stratum and pool
```{r}
# Summary
summary_stratumpool <- sites_upperlower %>% group_by(pool, stratum) %>%
  dplyr::summarize(npoints = n(),
            wood = sum(snag == 1),
            nowood = sum(snag == 0),
            propwood = wood/npoints) %>% as.data.frame()
summary_stratumpool

# Test for significant differences
chisq.test(summary_stratumpool[,c("wood", "nowood")])

# Confidence intervals
cis <- BinomCI(summary_stratumpool$wood, n = summary_stratumpool$npoints, conf.level = 0.95) %>% as.data.frame()
summary_stratumpool <- summary_stratumpool %>% mutate(lwr.ci = cis$lwr.ci,
                                                      upr.ci = cis$upr.ci)
row.names(summary_stratumpool) <- NULL

# Plot
summary_stratumpool %>% ggplot(aes(x = stratum, y = propwood, color = pool))+
  geom_point(size = 3,
             position = position_dodge(width = 0.3))+
  geom_errorbar(aes(ymin = lwr.ci, ymax = upr.ci),
                width = 0, size = 1.2, position = position_dodge(width = 0.3))+
  theme_minimal()+
  theme(text = element_text(size = 16))+
  ylab("Proportion of points with wood")+
  xlab("Habitat Stratum")
```

## By pool, year, and stratum
```{r}
# Summary
summary_poolyearstratum <- sites_upperlower %>% 
  group_by(pool, year, stratum) %>% 
  dplyr::summarize(npoints = n(),
            wood = sum(snag == 1),
            nowood = sum(snag == 0),
            propwood = wood/npoints) %>%
  as.data.frame()
head(summary_poolyearstratum)   

# Plot of proportions by stratum and pool
summary_poolyearstratum %>% ggplot(aes(x = year, y = propwood, group = stratum))+
  geom_line(aes(linetype = stratum))+
  ylim(0,1)+
  theme_minimal()+
  xlab("Year")+
  ylab("Proportion of points with wood")+
  facet_wrap(~pool)

# Plot of proportions by stratum and pool, only the two most common strata
summary_poolyearstratum %>% filter(stratum %in% c("MCB-U", "SCB")) %>%
  ggplot(aes(x = year, y = propwood, group = stratum))+
  geom_line(size = .65, aes(linetype = stratum))+
  theme_minimal()+
  xlab("Year")+
  ylab("Proportion of points with wood")+
  facet_wrap(~pool)

#Violin plots of annual variation by stratum and pool 
# All observations, by stratum and pool
summary_poolyearstratum %>% ggplot(aes(x = stratum, y = propwood))+
  geom_violin()+
  geom_boxplot(width = 0.1,
               fill = "black",
               outlier.color = NA)+
  stat_summary(fun.y = median,
               geom = "point",
               fill = "white",
               shape = 21,
               size = 2.5)+
  facet_wrap(~pool)

# Pools, including those with > 9 observations
summary_poolyearstratum %>% filter(npoints > 9) %>%
  ggplot(aes(x = pool, y = propwood))+
  geom_violin()+
  geom_boxplot(width = 0.1,
               fill = "black",
               outlier.color = NA)+
  stat_summary(fun.y = median,
               geom = "point",
               fill = "white",
               shape = 21,
               size = 2.5)

# Stratum and pool, greater than 10 observations
summary_poolyearstratum %>% filter(npoints > 9) %>%
  ggplot(aes(x = stratum, y = propwood))+
  geom_violin()+
  geom_boxplot(width=0.1,
               fill="black",
               outlier.colour=NA)+
  stat_summary(fun.y=median, 
               geom="point",
               fill="white",
               shape=21,
               size=2.5)+
  facet_wrap(~pool)
```

### Save summaries
```{r}
save(summary_poolyearstratum, file = "../data/outputs/summary_poolyearstratum.Rda")
save(summary_all, file = "../data/outputs/summary_all.Rda")
save(summary_pool, file = "../data/outputs/summary_pool.Rda")
save(summary_yearpool, file = "../data/outputs/summary_yearpool.Rda")
save(summary_stratum, file = "../data/outputs/summary_stratum.Rda")
save(summary_stratumpool, file = "../data/outputs/summary_stratumpool.Rda")
```
