---
title: "timeseries"
author: "Kaija Gahm"
date: "2/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup
## Libraries
```{r echo = FALSE}
source("libraries.R")
```

## Data
```{r}
load("../data/outputs/summary_all.Rda")
load("../data/outputs/summary_pool.Rda")
load("../data/outputs/summary_poolyearstratum.Rda")
load("../data/outputs/summary_stratum.Rda")
load("../data/outputs/summary_stratumpool.Rda")
load("../data/outputs/summary_yearpool.Rda")

summary_pool
```

# Time series
## Overall
### Examine by pool
```{r}
summary_yearpool %>% ggplot(aes(x = as.numeric(year), y = propwood))+
  geom_point()+
  stat_smooth(method = "loess", col = "red", se = F)+
  facet_wrap(~pool)+
  xlab("Year")+
  ylab("Proportion with LW")
```

### Tests on each pool
```{r}
head(summary_yearpool)
summary_yearpool <- summary_yearpool %>% mutate(year = as.numeric(year))

stats <- summary_yearpool %>% group_by(pool) %>% 
  summarize(changepoint = pettitt.test(propwood)$estimate[1],
            changepoint2 = pettitt.test(propwood)$estimate[2],
            u_stat = pettitt.test(propwood)$statistic,
            pt_pval = round(pettitt.test(propwood)$p.value, 3),
            changeyear = year[changepoint] + 0.5,
            changeyear2 = year[changepoint2] + 0.5,
            mk_z = round(mk.test(propwood)$statistic, 3),
            mk_pval = round(mk.test(propwood)$p.value, 3),
            begyear = min(year, na.rm = T),
            endyear = max(year, na.rm = T),
            mn1 = round(mean(propwood[1:changepoint]), 3),
            mn2 = round(mean(propwood[(changepoint + 1):n()]), 3),
            mn_overall = round(mean(propwood), 3),
            slope_coef = coef(lm(propwood ~ year))[2],
            intercept_coef = coef(lm(propwood ~ year))[1])
stats

# this table shows results of the pettitt test and the mk test for the _overall_ dataset. It does not yet include trend analyses for subsections if a breakpoint was found.

# I have added columns for multiple changepoints. Need to add stats for trend analyses for subsections. 
```

### Add regression coefficients to stats
```{r}
stats <- stats %>%
  mutate(slope_1 = NA,
         intercept_1 = NA,
         slope_2 = NA,
         intercept_2 = NA,
         mkp1 = NA,
         mkp2 = NA)
# Set stats vals to NA if not significant
for(i in 1:nrow(stats)){
  # pre- and post-changepoint regressions
  ## pre
  subdat_1 <- summary_yearpool %>% 
    filter(pool == stats$pool[i]) %>%  # select current pool
    filter(year >= stats$begyear[i] & year <= stats$changeyear[i]) # select years before the changepoint
  stats$slope_1[i] <- coef(lm(propwood ~ year, data = subdat_1))[2] # slope for pre-changepoint regression
  stats$intercept_1[i] <- coef(lm(propwood ~ year, data = subdat_1))[1] # intercept for pre-changepoint regression
  stats$mkp1[i] <- round(mk.test(subdat_1$propwood)$p.value, 3)
  
  ## post
  subdat_2 <- summary_yearpool %>% 
    filter(pool == stats$pool[i]) %>%  # select current pool
    filter(year >= stats$changeyear[i] & year <= stats$endyear[i]) # select years after the changepoint
  stats$slope_2[i] <- coef(lm(propwood ~ year, data = subdat_2))[2] # slope for post-changepoint regression
  stats$intercept_2[i] <- coef(lm(propwood ~ year, data = subdat_2))[1] # intercept for post-changepoint regression
  stats$mkp1[i] <- round(mk.test(subdat_1$propwood)$p.value, 3)
  stats$mkp2[i] <- round(mk.test(subdat_2$propwood)$p.value, 3)
  
  if(stats$pt_pval[i] >= 0.05){ # if the changepoint is non-significant...
    stats[i,c(2:4, 6:7, 12:13)] <- NA # set the changepoints, changeyears, and means to NA. 
  }else{
    stats[i, "mn_overall"] <- NA # if the p val is significant, set the overall mean to NA.
  }
  if(stats$mk_pval[i] >= 0.05){ # if the overall mk test is non-significant...
    stats[i, c("slope_coef", "intercept_coef")] <- NA # set the linear regression values to NA
  }
  if(stats$mkp1[i] >= 0.05){ # if the first-period mk test is non-significant...
    stats[i, c("slope_1", "intercept_1")] <- NA # set the first-period linear regression values to NA
  }
  if(stats$mkp2[i] >= 0.05){ # if the second-period mk test is non-significant...
    stats[i, c("slope_2", "intercept_2")] <- NA # set the second-period linear regression values to NA.
  } 
}

# Use the first and second linear regression values to compute y values at the ends of each regression segment
stats <- stats %>%
  mutate(ybeg_1 = slope_1*begyear + intercept_1,
         yend_1 = slope_1*changeyear + intercept_1,
         ybeg_2 = slope_2*changeyear + intercept_2,
         yend_2 = slope_2*endyear + intercept_2)
```


### Make plots (facetted)
```{r, warning = FALSE}
summary_yearpool %>%
  ggplot(aes(x = year, y = propwood))+
  geom_point()+
  facet_wrap(~pool)+
  theme_bw()+
  theme(panel.grid = element_blank(),
        text = element_text(size = 18))+
  ylab("Proportion of sites with large wood")+
  xlab("Year")+
  geom_vline(data = stats, aes(xintercept = changeyear), col = "red", lty = 2)+
  #geom_vline(data = stats, aes(xintercept = changeyear2), col = "red", lty = 2)+
  geom_abline(data = stats, aes(slope = slope_coef, intercept = intercept_coef),
              col = "black", lwd = 0.5)+ #overall regression
  geom_segment(data = stats, aes(x = begyear, y = mn1, xend = changeyear, yend = mn1), 
               col = "darkgrey", lty = 2)+ # mean 1
  geom_segment(data = stats, aes(x = changeyear, y = mn2, xend = endyear, yend = mn2),
               col = "darkgrey", lty = 2)+ # mean 2
  geom_segment(data = stats, aes(x = begyear, y = mn_overall, xend = endyear, yend = mn_overall),
               col = "darkgrey", lty = 2)+ # overall mean
  geom_segment(data = stats, aes(x = begyear, y = ybeg_1, xend = changeyear, yend = yend_1),
               col = "black", lwd = 1.5)+ # regression 1
  geom_segment(data = stats, aes(x = changeyear, y = ybeg_2, xend = endyear, yend = yend_2),
               col = "black", lwd = 1.5) # regression 2
```

## By aquatic areas
### Tests on each aquatic area within each pool
#### Calculate stats
```{r}
summary_poolyearstratum
summary_poolyearstratum <- summary_poolyearstratum %>% mutate(year = as.numeric(year))


stats_stratum <- summary_poolyearstratum %>% 
  group_by(pool, stratum) %>%
  filter(n() >= 3) %>%
  summarize(changepoint = pettitt.test(propwood)$estimate[1],
            changepoint2 = pettitt.test(propwood)$estimate[2],
            u_stat = pettitt.test(propwood)$statistic,
            pt_pval = round(pettitt.test(propwood)$p.value, 3),
            changeyear = year[changepoint] + 0.5,
            changeyear2 = year[changepoint2] + 0.5,
            mk_z = round(mk.test(propwood)$statistic, 3),
            mk_pval = round(mk.test(propwood)$p.value, 3),
            begyear = min(year, na.rm = T),
            endyear = max(year, na.rm = T),
            mn1 = round(mean(propwood[1:changepoint]), 3),
            mn2 = round(mean(propwood[(changepoint + 1):n()]), 3),
            mn_overall = round(mean(propwood), 3),
            slope_coef = coef(lm(propwood ~ year))[2],
            intercept_coef = coef(lm(propwood ~ year))[1])
  
stats_stratum
```

#### Edit stats
```{r}
stats_stratum <- stats_stratum %>%
  mutate(slope_1 = NA,
         intercept_1 = NA,
         slope_2 = NA,
         intercept_2 = NA,
         mkp1 = NA,
         mkp2 = NA,
         fstat_1 = NA,
         r2adj_1 = NA,
         fstat_2 = NA,
         r2adj_2 = NA)
i <- 1
head(subdat_1)
```

```{r}
for(i in 1:nrow(stats_stratum)){
  # pre- and post-changepoint regressions
  ## pre
  subdat_1 <- summary_poolyearstratum %>% 
    filter(pool == stats_stratum$pool[i]) %>%  # select current pool
    filter(year >= stats_stratum$begyear[i] & year <= stats_stratum$changeyear[i]) %>% # select years before the changepoint
    filter(stratum == stats_stratum$stratum[i])
  
  stats_stratum$slope_1[i] <- coef(lm(propwood ~ year, data = subdat_1))[2] # slope for pre-changepoint regression
  stats_stratum$intercept_1[i] <- coef(lm(propwood ~ year, data = subdat_1))[1] # intercept for pre-changepoint regression
  stats_stratum$fstat_1[i] <- summary(lm(propwood ~ year, data = subdat_1))$fstatistic[1]
# F-statistic for pre-changepoint regression
  stats_stratum$r2adj_1[i] <- summary(lm(propwood ~ year, data = subdat_1))$adj.r.squared # adjusted r-squared for pre-changepoint regression
  stats_stratum$mkp1[i] <- round(mk.test(subdat_1$propwood)$p.value, 3)
  
  ## post
  subdat_2 <- summary_poolyearstratum %>% 
    filter(pool == stats_stratum$pool[i]) %>%  # select current pool
    filter(year >= stats_stratum$changeyear[i] & year <= stats_stratum$endyear[i]) %>% # select years after the changepoint
    filter(stratum == stats_stratum$stratum[i])
  
  stats_stratum$slope_2[i] <- coef(lm(propwood ~ year, data = subdat_2))[2] # slope for post-changepoint regression
  stats_stratum$intercept_2[i] <- coef(lm(propwood ~ year, data = subdat_2))[1] # intercept for post-changepoint regression
  stats_stratum$fstat_2[i] <- summary(lm(propwood ~ year, data = subdat_2))$fstatistic[1] # F-statistic for post-changepoint regression
  stats_stratum$r2adj_2[i] <- summary(lm(propwood ~ year, data = subdat_2))$adj.r.squared # adjusted r-squared for pre-changepoint regression
  stats_stratum$mkp2[i] <- round(mk.test(subdat_2$propwood)$p.value, 3)
}


  # if(stats_stratum$pt_pval[i] >= 0.05){ # if the changepoint is non-significant...
  #   stats_stratum[i,c("changepoint", "changepoint2", "changeyear", "changeyear2", "mn1", "mn2", "slope_1", "intercept_1", "slope_2", "intercept_2", "mkp1", "mkp2")] <- NA # set the changepoints, changeyears, means, and everything relating to the first- and second-period regressions to NA 
  # }else{
  #   stats_stratum[i, "mn_overall"] <- NA # if the p val is significant, set the overall mean to NA.
  # }
  # if(stats_stratum$mk_pval[i] >= 0.05){ # if the overall mk test is non-significant...
  #   stats_stratum[i, c("slope_coef", "intercept_coef")] <- NA # set the linear regression values to NA
  # }
  # if(!is.na(stats_stratum$mkp1[i]) & stats_stratum$mkp1[i] >= 0.05){ # if the first-period mk test is non-significant...
  #   stats_stratum[i, c("slope_1", "intercept_1")] <- NA # set the first-period linear regression values to NA
  # }
  # if(!is.na(stats_stratum$mkp2[i]) & stats_stratum$mkp2[i] >= 0.05){ # if the second-period mk test is non-significant...
  #   stats_stratum[i, c("slope_2", "intercept_2")] <- NA # set the second-period linear regression values to NA.
  # } 



# Use the first and second linear regression values to compute y values at the ends of each regression segment
stats_stratum <- stats_stratum %>%
  mutate(ybeg_1 = slope_1*begyear + intercept_1,
         yend_1 = slope_1*changeyear + intercept_1,
         ybeg_2 = slope_2*changeyear + intercept_2,
         yend_2 = slope_2*endyear + intercept_2)


# View the stats needed for Table S4.
stats_stratum %>%
  filter(pt_pval <= 0.05) %>% # only significant breakpoints
  dplyr::select(pool, stratum, begyear, changeyear, fstat_1, fstat_2, mkp1, mkp2, r2adj_1, r2adj_2) %>%
  mutate(fstat_1 = case_when(mkp1 <= 0.05 ~ fstat_1), # only significant trend lines
         fstat_2 = case_when(mkp2 <= 0.05 ~ fstat_2),
         r2adj_1 = case_when(mkp1 <= 0.05 ~ r2adj_1),
         r2adj_2 = case_when(mkp2 <= 0.05 ~ r2adj_2))
```

### Make plots (facetted and colored)
```{r, warning = FALSE}
# Partial regressions only
summary_poolyearstratum %>%
  ggplot(aes(x = year, y = propwood, col = stratum))+
  geom_point(size = 1.5)+
  facet_wrap(~pool)+
  theme_bw()+
  theme(panel.grid = element_blank(),
        text = element_text(size = 18))+
  ylab("Proportion of sites with large wood")+
  xlab("Year")+
  geom_vline(data = stats_stratum, aes(xintercept = changeyear, col = stratum), lwd = 1, lty = 2, show.legend = F)+
  #geom_abline(data = stats_stratum, aes(slope = slope_coef, intercept = intercept_coef, col = stratum),
  #            lwd = 0.5, show.legend = F)+ #overall regression
  #geom_segment(data = stats_stratum, aes(x = begyear, y = mn1, xend = changeyear, yend = mn1, col = stratum), lty = 2)+ # mean 1 
  #geom_segment(data = stats_stratum, aes(x = changeyear, y = mn2, xend = endyear, yend = mn2, col = stratum), lty = 2)+ # mean 2 
  #geom_segment(data = stats_stratum, aes(x = begyear, y = mn_overall, xend = endyear, yend = mn_overall, col = stratum), lty = 2)+ # overall mean 
  geom_segment(data = stats_stratum, aes(x = begyear, y = ybeg_1, xend = changeyear, yend = yend_1, col = stratum), lwd = 1, show.legend = F)+ # regression 1
  geom_segment(data = stats_stratum, aes(x = changeyear, y = ybeg_2, xend = endyear, yend = yend_2, col = stratum), lwd = 1, show.legend = F)+ # regression 2
  guides(color = guide_legend(override.aes = list(size = 2)))+
  theme(legend.position = "bottom")+
  scale_color_viridis_d()+
  NULL

# Overall regressions only
summary_poolyearstratum %>%
  ggplot(aes(x = year, y = propwood, col = stratum))+
  geom_point(size = 1.5)+
  facet_wrap(~pool)+
  theme_bw()+
  theme(panel.grid = element_blank(),
        text = element_text(size = 18))+
  ylab("Proportion of sites with large wood")+
  xlab("Year")+
  geom_vline(data = stats_stratum, aes(xintercept = changeyear, col = stratum), lwd = 1, lty = 2, show.legend = F)+
  geom_abline(data = stats_stratum, aes(slope = slope_coef, intercept = intercept_coef, col = stratum),
              lwd = 1, show.legend = F)+ #overall regression
  #geom_segment(data = stats_stratum, aes(x = begyear, y = mn1, xend = changeyear, yend = mn1, col = stratum), lty = 2)+ # mean 1 
  #geom_segment(data = stats_stratum, aes(x = changeyear, y = mn2, xend = endyear, yend = mn2, col = stratum), lty = 2)+ # mean 2 
  #geom_segment(data = stats_stratum, aes(x = begyear, y = mn_overall, xend = endyear, yend = mn_overall, col = stratum), lty = 2)+ # overall mean 
  #geom_segment(data = stats_stratum, aes(x = begyear, y = ybeg_1, xend = changeyear, yend = yend_1, col = stratum), lwd = 0.5, show.legend = F)+ # regression 1
  #geom_segment(data = stats_stratum, aes(x = changeyear, y = ybeg_2, xend = endyear, yend = yend_2, col = stratum), lwd = 0.5, show.legend = F)+ # regression 2
  guides(color = guide_legend(override.aes = list(size = 2)))+
  theme(legend.position = "bottom")+
  scale_color_viridis_d()+
  NULL
```

### Make plots (black and white)
```{r, warning = FALSE}
# Trend lines for portions of time series only
summary_poolyearstratum %>%
  ggplot(aes(x = year, y = propwood))+
  geom_point(aes(shape = stratum))+
  facet_wrap(~pool)+
  theme_bw()+
  theme(panel.grid = element_blank(),
        text = element_text(size = 18))+
  ylab("Proportion of sites with large wood")+
  xlab("Year")+
  geom_vline(data = stats_stratum, aes(xintercept = changeyear, lty = stratum))+
  #geom_abline(data = stats_stratum, aes(slope = slope_coef, intercept = intercept_coef, lty = stratum), lwd = 0.5, show.legend = FALSE)+ #overall regression
  #geom_segment(data = stats_stratum, aes(x = begyear, y = mn1, xend = changeyear, yend = mn1, col = stratum), lty = 2)+ # mean 1 
  #geom_segment(data = stats_stratum, aes(x = changeyear, y = mn2, xend = endyear, yend = mn2, col = stratum), lty = 2)+ # mean 2 
  #geom_segment(data = stats_stratum, aes(x = begyear, y = mn_overall, xend = endyear, yend = mn_overall, col = stratum), lty = 2)+ 
  geom_segment(data = stats_stratum, aes(x = begyear, y = ybeg_1, xend = changeyear, yend = yend_1, lty = stratum), lwd = 0.5, show.legend = FALSE)+ # regression 1
  geom_segment(data = stats_stratum, aes(x = changeyear, y = ybeg_2, xend = endyear, yend = yend_2, lty = stratum), lwd = 0.5, show.legend = FALSE)+ # regression 2
  theme(legend.position = "bottom")+
  scale_linetype_manual(values = c("dashed", "dotted", "solid", "longdash"))+
  scale_shape_manual(values = c(3, 4, 19, 1))+
  NULL

# Trend lines for entire time series only
summary_poolyearstratum %>%
  ggplot(aes(x = year, y = propwood))+
  geom_point(aes(shape = stratum))+
  facet_wrap(~pool)+
  theme_bw()+
  theme(panel.grid = element_blank(),
        text = element_text(size = 18))+
  ylab("Proportion of sites with large wood")+
  xlab("Year")+
  geom_vline(data = stats_stratum, aes(xintercept = changeyear, lty = stratum))+
  geom_abline(data = stats_stratum, aes(slope = slope_coef, intercept = intercept_coef, lty = stratum), lwd = 0.5, show.legend = FALSE)+ #overall regression
  #geom_segment(data = stats_stratum, aes(x = begyear, y = mn1, xend = changeyear, yend = mn1, col = stratum), lty = 2)+ # mean 1 
  #geom_segment(data = stats_stratum, aes(x = changeyear, y = mn2, xend = endyear, yend = mn2, col = stratum), lty = 2)+ # mean 2 
  #geom_segment(data = stats_stratum, aes(x = begyear, y = mn_overall, xend = endyear, yend = mn_overall, col = stratum), lty = 2)+ 
  #geom_segment(data = stats_stratum, aes(x = begyear, y = ybeg_1, xend = changeyear, yend = yend_1, lty = stratum), lwd = 1, show.legend = FALSE)+ # regression 1
  #geom_segment(data = stats_stratum, aes(x = changeyear, y = ybeg_2, xend = endyear, yend = yend_2, lty = stratum), lwd = 1, show.legend = FALSE)+ # regression 2
  theme(legend.position = "bottom")+
  scale_linetype_manual(values = c("dashed", "dotted", "solid", "longdash"))+
  scale_shape_manual(values = c(3, 4, 19, 1))+
  NULL
```

# Exploring discharge data
```{r}
# load
dis <- read.csv("../data/inputs/pools_discharge_1975_2018.csv", as.is = T)

# clean up
dis <- dis %>%
  pivot_longer(., names_to = "pool", cols = c("P4", "P8", "P26", "P13", "OR", "LAG")) %>%
  mutate(pool = fct_recode(factor(pool),
                           "4" = "P4",
                           "8" = "P8",
                           "13" = "P13",
                           "26" = "P26",
                           "OR" = "OR",
                           "LG" = "LAG")) %>%
  rename(date = Dates) %>%
  mutate(date = mdy(date)) %>%
  mutate(year = year(date))

# aggregate by year
dis_year <- dis %>%
  group_by(pool, year) %>%
  summarize(meandis = mean(value, na.rm = T))

head(summary_yearpool)

# join wood data
dis_year <- dis_year %>%
  left_join(summary_yearpool, by = c("pool", "year")) %>%
  filter(!is.na(propwood))

head(dis_year)
```

### Relationship between mean discharge and proportion of sites with wood, broken down by pool
```{r}
dis_year %>%
  ggplot(aes(x = meandis, y = propwood))+
  geom_point()+
  facet_wrap(~pool, scales = "free") # not initially seeing any pattern, except maybe in OR.

dis_year %>%
  ggplot(aes(x = meandis, y = propwood))+
  geom_point()+
  facet_wrap(~pool, scales = "free") +
  geom_smooth(method = "lm") # Yeah, huh, that's kind of interesting. Essentially no pattern in any pool except OR. Maybe *slight* upward slopes in the other pools, but not at all significant.
```

### Discharge patterns over time
```{r}
# Just plot discharge over time
dis_year %>%
  ggplot(aes(x = year, y = meandis, col = pool))+
  geom_line() # this is kind of interesting just by itself. I guess it makes sense--as you go down the river, the discharge gets higher, but the magnitude of the differences also changes.
```

### Standardized discharge
```{r}
# What about looking at discharge in terms of deviation from the mean for each pool? I.e. do extreme years relate to proportion of wood observed?
dis_year <- dis_year %>%
  group_by(pool) %>%
  mutate(meandis_std = scale(meandis)) # scale centers the data around the mean

# std discharge vs. propwood
dis_year %>%
  ggplot(aes(x = meandis_std, y = propwood))+
  geom_point()+
  facet_wrap(~pool)+
  geom_smooth(method = "lm") # I guess this is kind of the same as what I did before, by just letting the x scales be free. Oops. What if I also standardize the y scales?

```

Standardized wood proportion
```{r}
# Standardizing the wood proportion
dis_year <- dis_year %>%
  group_by(pool) %>%
  mutate(propwood_std = scale(propwood))

# std discharge vs. std propwood
dis_year %>%
  ggplot(aes(x = meandis_std, y = propwood_std))+
  geom_point()+
  facet_wrap(~pool)+
  geom_smooth(method = "lm") # yeah, that didn't really change things too much. 
```

Both vs. time
```{r}
# What about standardized wood proportion and standardized discharge both vs. year?
dis_year %>%
  ggplot(aes(x = year, y = propwood_std))+ # proportion of wood is black; discharge is red.
  geom_line()+
  facet_wrap(~pool)+
  geom_line(aes(x = year, y = meandis_std), col = "red")
```

Wood vs. previous year's discharge?
```{r}
# add previous year's discharge
dis_year <- dis_year %>%
  group_by(pool) %>%
  mutate(lagdis_std = lag(meandis_std))

dis_year %>%
  ggplot(aes(x = lagdis_std, y = propwood_std))+
  geom_point()+
  facet_wrap(~pool)+
  geom_smooth(method = "lm") # meh, kind of inconclusive.
  
```







